### 2021-06-07 memo.tex
alternative approach

>>>>

\section{Via direct study of line operators}

Let's determine the anomaly/extension of so SQCD by studying the line operators.
Here I concentrate on the case of $SO(2n_c)_\pm$ with $2n_f$ flavors.

Let us first consider the case of pure $SO(2n_c)_\pm$ with $n_c$ odd,
and recall how to distinguish $\bZ_2\times \bZ_2$ or $\bZ_4$.

In either case, the group of the charges of the line operators has the structure \[
0\to \bZ_2 \to G \to \bZ_2 \to 0
\] where the $\bZ_2$ subgroup is generated by the 't Hooft line
and the $\bZ_2$ quotient is generated by the Wilson line in the vector representation.
To tell whether $G$ is $\bZ_2\times \bZ_2$ or $\bZ_4$,
we simply take two 't Hooft lines.
When combined, it is a charge-2 't Hooft line, which can be screened by the dynamical monopole of this charge.
In the $SO_+$ theory, this dynamical monopole is electrically neutral,
while the $SO_-$ theory, it has the electric charge equal to the vector representation.
Therefore, the charge-2 't Hooft line is equivalent to having no line operator in $SO_+$ theory,
while it is equivalent to having a Wilson line in the vector representation in $SO_-$.
We find that $G$ is $\bZ_2\times \bZ_2$ for $SO_+$ and $\bZ_4$ for $SO_-$.

We learned the important lesson: 
the 1-form symmetry group is $\bZ_2\times \bZ_2$ or $\bZ_4$ depending on 
whether the electric charge of the dynamical monopole of an appropriate charge is 
neutral or vector.

So we would like to determine that in the $SO$ SQCD,
but the direct analysis of dynamical monopoles in this theory is hard.
For this purpose, we make the following deformations which do not change the symmetry structure:
\begin{itemize}
\item We reduce the flavor symmetry from $su(2n_f)$ to $usp(2n_f)$.
\item We add an adjoint scalar $\Phi_{[ab]}$ and the interaction $\psi^{ai}_\alpha \psi^{bj}_\beta J_{ij} \Phi_{ab}\epsilon^{\alpha\beta} +cc$, where $J_{[ij]}$ is the constant matrix for the $usp(2n_f)$ part.
\item We give a vev to $\Phi_{ab}$ to break $so(2n_c)$ to $so(2)^{n_c}$.
\end{itemize}
The 't Hooft line of the original $SO(2n_c)$ theory is in the `vector' class
$(1,0,\ldots,0)$ under $so(2)^{n_c}$,
and therefore the dynamical monopole we need to analyze has the charge $(2,0,\ldots,0)$.

Since the fermions are in the vector representation of $so(2n_c)$,
they are charge 2 under the $so(2)$, and should give rise to fermionic zero modes transforming in \begin{equation}
V_{2} \otimes R_{2n_f} 
\end{equation} where $V_2$ is the doublet of spacetime $so(3)$, 
$R_{2n_f}$ is the fundamental of $usp(2n_f)$,
and we need to impose the reality condition using the pseudoreality of both factors,
so that there are $4n_f$ Majorana fermion in total.

I haven't analyzed the general case; let's take $n_f=1$ and $n_f=2$ as examples.
When $n_f=1$, there are $4$ Majorana fermions.
Quantizing them, we find the monopoles in \begin{equation}
V_2 \otimes \mathbf{1} \oplus \mathbf{1}\otimes R_2.
\end{equation}
I think we should take the spacetime scalar as indicative, (which needs to be justified)
so it has the `vector' charge under $usp(2)$ flavor symmetry.
Therefore the 1-form symmetry is extended.

When $n_f=2$, there are $8$ Majorana fermions, 
and spatial $su(2)$ and flavor $usp(4)$ are embedded into $so(8)$ via \begin{equation}
su(2)\times usp(4) \simeq so(3)\times so(5) \subset so(8).
\end{equation}
Therefore, the monopoles are easily seen to be in \begin{equation}
V_3\otimes \mathbf{1} \oplus V_2\otimes R_2 \oplus V_1\otimes R_5.
\end{equation}
Again taking the spacetime scalar part, it is in the `adjoint' charge under the $usp(4)$ flavor symmetry.
Therefore the 1-form symmetry is not extended.

So far we considered $SO_+$ theories. 
In $SO_-$ theories, we need to add electric charges coming from the discrete theta angle.
This is trivial if $n_c$ is even, but is the `vector' charge if $n_c$ is odd.
